---
title: "Lecture 8"
author: "Esther Kim"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)
```

#rnoaa = user package
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"), #download these specific channels
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>% #time range you want data for
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", #renmaing each station ID 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>% #tenths of a degree celsius - divded by 10 to get whole number
  select(name, id, everything()) #want name and id to be listed first and everything else after

weather_df
```

#Scatterplot

miniminum temperature = x-axis
maximum temperature = y-axis

aes() = defining aesthetic mapping
- telling R what to put on x axis by putting `x =` and `y =`
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax))
```

Need to add geometry to the scatterplot above to get an output
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```
Warning: Removed 15 rows containing missing values (geom_point)
- being reminded by R that there were missing values and they were not included when the graph was created

`ggplot`
Can use ` %>% `
- This is useful if you want to perform a few data processing steps (e.g. mutating, filtering, dropping, etc.)

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()

weather_df %>% 
  drop_na() %>% 
  filter (name == "CentralPark_NY") %>% 
  ggplot(aes(x=tmin, y=tmax)) + geom_point()
```


Same plot but different
weather_scatterplot %>% 
  weather_df %>% 
  drop_na() %>% 
  ggplot(aes(x=tmin, y=tmax)) #base figure without any geometry imposed on it
```{r}
plot_weather = 
  weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) 

plot_weather + geom_point()
```

Adding color to distinguish different variables 
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name)) #name is variable 
```
Can define aesthetic mapping for specific geometries

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  geom_smooth(se = FALSE)
```
Different geometries can be stacked and be assigned their own aesthetic

`geom_smooth()` = takes all x and y data and runs a line/curve through dataset
- used to describe association
- useful in parts of exploratory association (e.g. is this a linear association? is there an association I care about?)

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name))+
  geom_point() +
  geom_smooth()
```
By adding the aesthetic to the first line (within `ggplot`), the geometries are applied to the entire dataset
- therefore, there will be three lines instead of one (the aesthetic is only within `geom_smooth()`)

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```
`se = FALSE` = removing the confidence intervals
`alpha = .5` = sets transparency

```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name) #defines column variable

ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(name ~ .) #defines row variable

ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .3) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```
`color =` = color gradiation
`size =` = size of each point differs 
- using size is useful when it is difficult to distinguish the different color points on the graph

Rotating axis
```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .3) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name) +
  theme(axis.text.x = element_text (angle = 90, vjust = 0.5, hjust = 1))
```

#Graphs
Histogram = use fill instead of color 
- color just outlines the bars
```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, fill = name)) + 
  geom_histogram() +
  facet_grid(. ~ name)
```
use `facet_grid` to make it easier to visualize each bar in the graph
- without it, the graphs lay on top of one another and may be difficult to interpret

Write a code chain that starts with `weather_df`; focuses only on Central Park, converts temperatures to Fahrenheit, makes a scatterplot of min vs. max temperature, and overlays a linear regression line (using options in `geom_smooth())`.
```{r}
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  mutate(
    tmax_fahr = tmax * (9 / 5) + 32,
    tmin_fahr = tmin * (9 / 5) + 32) %>% 
  ggplot(aes(x = tmin_fahr, y = tmax_fahr)) +
  geom_point(alpha = .5) + 
  geom_smooth(method = "lm", se = FALSE)

```

```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_smooth(se = FALSE) 
```

```{r}
ggplot(weather_df, aes(x = tmax, y = tmin)) + 
  geom_hex()
```

In the preceding, we set the alpha aesthetic “by hand” instead of mapping it to a variable. This is possible for other aesthetics too. To check your understanding of this point, try to explain why the two lines below don’t produce the same result:
```{r}
ggplot(weather_df) + geom_point(aes(x = tmax, y = tmin), color = "blue")
ggplot(weather_df) + geom_point(aes(x = tmax, y = tmin, color = "blue"))
```
In the first attempt, we’re defining the color of the points by hand; in the second attempt, we’re implicitly creating a color variable that has the value `blue` everywhere; `ggplot` is then assigning colors according to this variable using the default color scheme.


```{r}
ggplot(weather_df, aes(x = tmax)) + 
  geom_histogram()
```

```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_histogram(position = "dodge", binwidth = 2)
```

```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue")
```

```{r}
ggplot(weather_df, aes(x = name, y = tmax)) + geom_boxplot()
```

```{r}
ggplot(weather_df, aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), alpha = .5) + 
  stat_summary(fun = "median", color = "blue")
```

```{r}
ggplot(weather_df, aes(x = tmax, y = name)) + 
  geom_density_ridges(scale = .85)
```

First a density plot:
```{r}
ggplot(weather_df, aes(x = prcp)) + 
  geom_density(aes(fill = name), alpha = .5) 
```

Next a ridge plot:
```{r}
weather_df %>% 
ggplot(aes(x = tmax, y = name)) + 
  geom_density_ridges(scale = .85)
```
when you have more than 5 plots, boxplots, ridge plots, and violin plots are more interpretable


Last a boxplot:
```{r}
ggplot(weather_df, aes(y = prcp, x = name)) + 
  geom_boxplot() 
```

Violin plots - useful when comparing a large number of plots (>10)
```{r}
weather_df %>% 
ggplot(aes(x = name, y = tmax, fill = name)) + 
  geom_violin()
```

```{r}
weather_plot = ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) 

ggsave("weather_plot.pdf", weather_plot, width = 8, height = 5)
```

Used to adjust how the figures are printed when `knit`
```{r, fig.width = 10}
weather_scatterplot()
```


Set these options instead of definingin in each `{r}` inline
```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name))
```



