---
title: "Lecture 9"
output: github_document
date: "2022-10-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

Conduct exploratory analyses using `dplyr` verbs (`group_by` and `summarize`), along with numeric data summaries.

# Example
We’ll continue in the same Git repo / R project that we used for visualization, and use essentially the same `weather_df` dataset – the only exception is the addition of `month` variable, created using `lubridate::floor_date()`.

```{r}
weather_df =  
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())
```

## `group_by`

Datasets are often comprised of groups defined by one or more (categorical) variable; `group_by()` makes these groupings explicit so that they can be included in subsequent operations. For example, we might group `weather_df` by `name` and `month`:
```{r}
weather_df %>%
  group_by(name, month)
```
Output: `Groups:   name, month [36]`
- 36 distinct groups: 12 groups for 3 names

Not changing the dataset in anyway

Several important functions respect grouping structures. You will frequently use summarize to create one-number summaries within each group, or use mutate to define variables within groups. The rest of this example shows these functions in action.

Because these (and other) functions will use grouping information if it exists, it is sometimes necessary to remove groups using `ungroup()`.



## *Counting things*
As an intro to `summarize`, let’s count the number of observations in each month in the complete `weather_df` dataset.
```{r}
weather_df %>%
  group_by(month) %>%
  summarize(n_obs = n())
```
For each month, how many observations is the output


```{r}
weather_df %>% 
  group_by(name) %>% 
  summarize(
    n_obs = n()
  )
```
`n()` = will count how many observations within each variable identified in `group_by()`

The output states that there are 3 groups and how many observations are in each group
- CentralPark_NY = 365
- Waikiki_HA = 365
- Waterhole_WA = 365


We can group by more than one variable, too.
```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(n_obs = n())
```
Will output the number of observation for each month of each group
- there will be 36 rows: 12 months for 3 groups

# `Count()`

```{r}
weather_df %>% 
  count(name)
```
Will do the same thing as `summarize(n_obs = n())`

Don't use `table()`!
```{r}
weather_df %>% 
  pull(month)
```
`pull()` similar to `$` - trying to pull a single variable or column from dataframe

```{r}
weather_df %>% 
  group_by(month) %>% 
  summarize(
    n_obs = n(),
    n_dist = n_distinct(month),
    n_dates = n_distinct(date)
  )
```
`n_distinct()` = count the number of unique values in a dataset

# 2X2 tables
```{r}
weather_df %>% 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ ""
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  janitor::tabyl(name, cold)
```
`janitor::tabyl` = creates 2X2 table

`case_when()`
- `TRUE      ~ ""` = included anything past the second line, making sure that there are no false left - any leftover values is labeled "" (can also use "THIS IS A PROBLEM" or something that helps flag the observations that do not satisfy the two prior logic statements)

## General Summaries
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    n_obs = n(),
    mean_tmax = mean(tmax, na.rm = TRUE),
    sd_prcp = sd(prcp),
    median_tmax = median(tmax)
  )
```
`mean_tmax = mean(tmax)` = creating a variable that represents the mean of variable `t_max`
`sd()` = standard deviation
`na.rm = TRUE` = remove N/A when computing
`median()` = median

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(across(prcp:tmin, mean))
```
`across(prcp:tmin,mean)` = taking the mean of all the variables from prcp to tmin
`across()` = applying functions across multiple columns

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE)
  ) %>% 
  ggplot(aes(x=month, y=mean_tmax, color = name))+
  geom_point()+
  geom_path()
```
`geom_path()` = connects the observations in the ggplot

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE)
  ) %>% 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax
  ) %>% 
  knitr::kable(digits = 2)
```
`pivot_wider()` = usually means the table is being "untidied"
- however, may be easier to interpret

`knitr::kable()` = Rmarkdown's table
- make a neater table to print in Rmarkdown
- nice format for table

lagged observations
Offsets, especially lags, are used to compare an observation to it’s previous value. This is useful, for example, to find the day-by-day change in max temperature within each station over the year.

`min_rank()` = ranking functions